---
title: "Deer Risk effects"
author: "Fiona Seaton"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.dim = c(7,5))
library(dplyr)
library(brms)
library(emmeans)
library(tidybayes)
library(ggplot2)
theme_set(theme_classic())
teal <- "#0383a4"
two_cols <- c("#999999","#0072B2","#E69F00")
model_loc <- "Outputs/Models/Deer/"
```

```{r, include = FALSE}
source("Helper_functions.R")
```

```{r deer risk data read, cache = TRUE}
deerrisk <- read.csv("Metadata/DEER_RISK.csv") %>%
  mutate(DEER = ifelse(DEER == "None", "Low", DEER))
```



```{r data read}
BRLEAF_SUMMARY <- read.csv("Outputs/Broadleaf_summary_metrics.csv")
BRLEAF_SUMMARY <- BRLEAF_SUMMARY %>%
  inner_join(deerrisk, by = c("SITE_NO" = "SITE_NUMBER")) %>%
  mutate(YR = c("1971","2001","2022")[YEAR],
         YDAY = (DAYOFYEAR - 200)/30,
         SITE = as.character(SITE_NO),
         PLOT = paste(SITE_NO, PLOT_NO, sep = "_"),
         DEER = factor(DEER, c("Low","Moderate","High"),
                       ordered = TRUE))
```


# Check priors for count models



```{r count model sample prior}
mod_pr <- prior(normal(2,1), class = "b", coef = "YR1971") +
  prior(normal(2,1), class = "b", coef = "YR2001") +
  prior(normal(2,1), class = "b", coef = "YR2022") +
  prior(normal(0,1), class = "b", coef = "YDAY") +
  prior(normal(0,1), class = "b", coef = "moDEER") +
  prior(normal(0,1), class = "b", coef = "moDEER:YR2001") +
  prior(normal(0,1), class = "b", coef = "moDEER:YR2022") +
  prior(dirichlet(1), class = "simo", coef = "moDEER1") +
  prior(dirichlet(1), class = "simo", coef = "moDEER:YR20011") +
  prior(dirichlet(1), class = "simo", coef = "moDEER:YR20221") +
  prior(student_t(5, 0, 1), class = "sd")
```

```{r count check prior predictions, cache  = TRUE}
test_mod <- brm(SPECIES_RICH ~ -1 + YR*mo(DEER) + YDAY + (1|SITE/PLOT),
                data = BRLEAF_SUMMARY, family = "poisson",
                prior = mod_pr, cores = 4, sample_prior = "only")
plot(test_mod)
pp_check(test_mod)
pp_check(test_mod, "ecdf_overlay", ndraws = 20) +
  scale_x_continuous(limits = c(0,200))
```


### Species richness

```{r species richness model final}
rich_mod <- brm(SPECIES_RICH ~ -1 + YR*mo(DEER) + YDAY + (1|SITE/PLOT),
                data = BRLEAF_SUMMARY, family = "negbinomial",
                prior = mod_pr, cores = 4, warmup = 2000, iter = 6000, thin = 4,
                file = paste0(model_loc, "RICHNESS_NB"))
par_summary(rich_mod)
```

```{r richness model plots}
mcmc_plot(rich_mod, type = "rank_overlay")
pp_check(rich_mod, type = "ecdf_overlay", ndraws = 20)
resid_bysite(rich_mod)
```

Model fit seems good, no problems with Rhat, ESS or the fit to the data.

```{r emmeans richness pairwise comparison table}
rich_emm <- emmeans(rich_mod, ~ DEER | YR)
pairs(rich_emm) %>%
  knitr::kable(digits = 3, caption = "Richness on log-scale")
```

```{r emmeans richness pairwise comparison table v2}
pairs(rich_emm, type = "response") %>%
  knitr::kable(digits = 3, caption = "Richness on response-scale")
```

```{r emmeans richness plot}
rich_emm %>%
  gather_emmeans_draws() %>%
  mutate(Year = as.numeric(as.character(YR)),
         .value = exp(.value)) %>%
  ggplot(aes(x = Year, y = .value, colour = DEER, fill = DEER)) +
  stat_lineribbon(alpha = 1/4) +
  scale_fill_manual(values = two_cols,
                    aesthetics = c("colour","fill"),
                    name = "Deer Risk") +
  theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14)) +
  labs(y = "Species Richness")
```



### Stem Count

```{r stem model}
stem_mod <- brm(STEM_COUNT ~ -1 + YR*mo(DEER) + YDAY + (1|SITE/PLOT),
                data = BRLEAF_SUMMARY, family = "negbinomial",
                prior = mod_pr, cores = 4, warmup = 2000, iter = 6000, thin = 4,
                file = paste0(model_loc, "STEM_COUNT_NB"))
par_summary(stem_mod)
```

```{r stem model plots}
mcmc_plot(stem_mod, type = "rank_overlay")
pp_check(stem_mod, type = "ecdf_overlay", ndraws = 20)
resid_bysite(stem_mod)
```


```{r emmeans stempairwise comparison table}
stem_emm <- emmeans(stem_mod, ~ DEER | YR)
pairs(stem_emm) %>%
  knitr::kable(digits = 3, caption = "Stem Count on log-scale")
```

```{r emmeans stem pairwise comparison table v2}
pairs(stem_emm, type = "response") %>%
  knitr::kable(digits = 3, caption = "Stem Count on response-scale")
```

```{r emmeans stem plot}
stem_emm %>%
  gather_emmeans_draws() %>%
  mutate(Year = as.numeric(as.character(YR)),
         .value = exp(.value)) %>%
  ggplot(aes(x = Year, y = .value, colour = DEER, fill = DEER)) +
  stat_lineribbon(alpha = 1/4) +
  scale_fill_manual(values = two_cols,
                    aesthetics = c("colour","fill"),
                    name = "Deer Risk") +
  theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14)) +
  labs(y = "Stem Count")
```

### Stem Richness

```{r treerich model}
treerich_mod <- brm(TREE_RICHNESS ~ -1 + YR*mo(DEER) + YDAY + (1|SITE/PLOT),
                    data = BRLEAF_SUMMARY, family = "poisson",
                    prior = mod_pr, cores = 4, warmup = 2000, iter = 6000, thin = 4,
                    control = list(adapt_delta = 0.95),
                    file = paste0(model_loc, "TREE_RICHNESS"))
par_summary(treerich_mod)
```

```{r treerich model plots}
mcmc_plot(treerich_mod, type = "rank_overlay")
pp_check(treerich_mod, type = "ecdf_overlay", ndraws = 20)
resid_bysite(treerich_mod)
```


```{r emmeans treerichpairwise comparison table}
treerich_emm <- emmeans(treerich_mod, ~ DEER | YR)
pairs(treerich_emm) %>%
  knitr::kable(digits = 3, caption = "Tree Richness on log-scale")
```

```{r emmeans treerich pairwise comparison table v2}
pairs(treerich_emm, type = "response") %>%
  knitr::kable(digits = 3, caption = "Tree Richness on response-scale")
```

```{r emmeans treerich plot}
treerich_emm %>%
  gather_emmeans_draws() %>%
  mutate(Year = as.numeric(as.character(YR)),
         .value = exp(.value)) %>%
  ggplot(aes(x = Year, y = .value, colour = DEER, fill = DEER)) +
  stat_lineribbon(alpha = 1/4) +
  scale_fill_manual(values = two_cols,
                    aesthetics = c("colour","fill"),
                    name = "Deer Risk") +
  theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14)) +
  labs(y = "Tree Richness")
```




# Gaussian distributed responses

### Check priors for gaussian models


```{r gauss model sample prior}
mod_pr <- prior(normal(6,2), class = "b", coef = "YR1971") +
  prior(normal(6,2), class = "b", coef = "YR2001") +
  prior(normal(6,2), class = "b", coef = "YR2022") +
  prior(normal(0,1), class = "b", coef = "YDAY") +
  prior(normal(0,1), class = "b", coef = "moDEER") +
  prior(normal(0,1), class = "b", coef = "moDEER:YR2001") +
  prior(normal(0,1), class = "b", coef = "moDEER:YR2022") +
  prior(dirichlet(1), class = "simo", coef = "moDEER1") +
  prior(dirichlet(1), class = "simo", coef = "moDEER:YR20011") +
  prior(dirichlet(1), class = "simo", coef = "moDEER:YR20221") +
  prior(student_t(5, 0, 3), class = "sd") +
  prior(student_t(5, 0, 3), class = "sigma")
```

```{r gauss check prior predictions, cache = TRUE}
test_mod <- brm(SPECIES_SHANNON ~ -1 + YR*mo(DEER) + YDAY + (1|SITE/PLOT),
                data = BRLEAF_SUMMARY, family = "gaussian",
                prior = mod_pr, cores = 4, sample_prior = "only")
plot(test_mod, ask = FALSE)
pp_check(test_mod)
pp_check(test_mod, "ecdf_overlay", ndraws = 20)
```


### Species diversity - Shannon

```{r species diversity model}
shan_mod <- brm(SPECIES_SHANNON ~ -1 + YR*mo(DEER) + YDAY + (1|SITE/PLOT),
                data = BRLEAF_SUMMARY, family = "gaussian",
                prior = mod_pr, cores = 4, warmup = 2000, iter = 6000, thin = 4,
                file = paste0(model_loc, "SPECIES_SHANNON"))
par_summary(shan_mod)
```

```{r species diversity model plots}
mcmc_plot(shan_mod, type = "rank_overlay")
pp_check(shan_mod, type = "ecdf_overlay", ndraws = 20)
resid_bysite(shan_mod)
```

```{r emmeans shannon pairwise comparison table}
shan_emm <- emmeans(shan_mod, ~ DEER | YR)
pairs(shan_emm) %>%
  knitr::kable(digits = 3)
```

```{r emmeans shannon plot}
shan_emm %>%
  gather_emmeans_draws() %>%
  mutate(Year = as.numeric(as.character(YR))) %>%
  ggplot(aes(x = Year, y = .value, colour = DEER, fill = DEER)) +
  stat_lineribbon(alpha = 1/4) +
  scale_fill_manual(values = two_cols,
                    aesthetics = c("colour","fill"),
                    name = "Deer Risk") +
  theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14)) +
  labs(y = "Shannon diversity")
```




# Constrained positive model

### Mean basal area
```{r}
mod_pr <- prior(normal(2,1), class = "b", coef = "YR1971") +
  prior(normal(2,1), class = "b", coef = "YR2001") +
  prior(normal(2,1), class = "b", coef = "YR2022") +
  prior(normal(0,1), class = "b", coef = "YDAY") +
  prior(normal(0,1), class = "b", coef = "moDEER") +
  prior(normal(0,1), class = "b", coef = "moDEER:YR2001") +
  prior(normal(0,1), class = "b", coef = "moDEER:YR2022") +
  prior(dirichlet(1), class = "simo", coef = "moDEER1") +
  prior(dirichlet(1), class = "simo", coef = "moDEER:YR20011") +
  prior(dirichlet(1), class = "simo", coef = "moDEER:YR20221") +
  prior(student_t(5, 0, 1), class = "sd") +
  prior(gamma(0.1,0.1), class = "shape") 
```

```{r}
basal_mod <- brm(MEAN_BASAL_AREA ~ -1 + YR*mo(DEER) + YDAY  + (1|SITE/PLOT),
                 data = BRLEAF_SUMMARY, family = "gamma",
                 prior = mod_pr, cores = 4,
                 file = paste0(model_loc, "MEAN_BASAL_AREA"))
par_summary(basal_mod)
```

```{r basal model plots}
mcmc_plot(basal_mod, type = "rank_overlay")
pp_check(basal_mod, type = "ecdf_overlay", ndraws = 20) +
  scale_x_continuous(limits = c(0,5000))
resid_bysite(basal_mod)
```

```{r emmeans basal pairwise comparison table}
basal_emm <- emmeans(basal_mod, ~ DEER | YR)
pairs(basal_emm) %>%
  knitr::kable(digits = 3, caption = "Mean basal area on log-scale")
```

```{r emmeans basal pairwise comparison table v2}
pairs(basal_emm, type = "response") %>%
  knitr::kable(digits = 3, caption = "Mean basal area on response-scale")
```

```{r emmeans basal plot}
basal_emm %>%
  gather_emmeans_draws() %>%
  mutate(Year = as.numeric(as.character(YR)),
         .value = exp(.value)) %>%
  ggplot(aes(x = Year, y = .value, colour = DEER, fill = DEER)) +
  stat_lineribbon(alpha = 1/4) +
  scale_fill_manual(values = two_cols,
                    aesthetics = c("colour","fill"),
                    name = "Deer Risk") +
  theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14)) +
  labs(y = bquote("Mean basal area (cm"^2*")"))
```




# Cover models


## Check priors for cover/gamma models


```{r prop model sample prior}
mod_pr <- prior(normal(2,1), class = "b", coef = "YR1971") +
  prior(normal(2,1), class = "b", coef = "YR2001") +
  prior(normal(2,1), class = "b", coef = "YR2022") +
  prior(normal(0,1), class = "b", coef = "YDAY") +
  prior(normal(0,1), class = "b", coef = "moDEER") +
  prior(normal(0,1), class = "b", coef = "moDEER:YR2001") +
  prior(normal(0,1), class = "b", coef = "moDEER:YR2022") +
  prior(dirichlet(1), class = "simo", coef = "moDEER1") +
  prior(dirichlet(1), class = "simo", coef = "moDEER:YR20011") +
  prior(dirichlet(1), class = "simo", coef = "moDEER:YR20221") +
  prior(student_t(5, 0, 1), class = "sd") +
  prior(gamma(0.1,0.1), class = "shape") +
  prior(normal(0,1), class = "b", dpar = "hu") +
  prior(normal(-1,1), class = "b", coef = "YR1971", dpar = "hu") +
  prior(normal(-1,1), class = "b", coef = "YR2001", dpar = "hu") +
  prior(normal(-1,1), class = "b", coef = "YR2022", dpar = "hu") +
  prior(student_t(2, 0, 1), class = "sd", dpar = "hu")
```

```{r prop check prior predictions, cache = TRUE}
test_mod <- brm(bf(FORB_COVER ~ -1 + YR*mo(DEER) + YDAY + (1|SITE/PLOT),
                   hu ~ -1 + YR*mo(DEER) + YDAY + (1|SITE/PLOT)),
                data = BRLEAF_SUMMARY, family = "hurdle_gamma",
                prior = mod_pr, cores = 4, sample_prior = "only")
plot(test_mod)
pp_check(test_mod)
pp_check(test_mod, "ecdf_overlay", ndraws = 20)
```

### Forb cover

```{r forb model final}
forb_mod <- brm(bf(FORB_COVER ~ -1 + YR*mo(DEER) + YDAY + (1|SITE/PLOT),
                   hu ~ -1 + YR*mo(DEER) + YDAY + (1|SITE/PLOT)),
                data = BRLEAF_SUMMARY, family = "hurdle_gamma",
                prior = mod_pr, cores = 4, warmup = 2000, iter = 6000, thin = 4,
                file = paste0(model_loc, "FORB_COVER"),
                control = list(adapt_delta = 0.9))
par_summary(forb_mod)
```

```{r forb model diag plot, fig.dim = c(7,8)}
mcmc_plot(forb_mod, type = "rank_overlay")
pp_check(forb_mod, type = "ecdf_overlay", ndraws = 20) +
  scale_x_continuous(limits = c(0,250))
resid_bysite(forb_mod)
```

```{r emmeans forb pairwise comparison table}
forb_emm <- emmeans(forb_mod, ~ DEER | YR, epred = TRUE)
pairs(forb_emm) %>%
  knitr::kable(digits = 3, caption = "Forb cover")
```


```{r emmeans forb plot}
forb_emm %>%
  gather_emmeans_draws() %>%
  mutate(Year = as.numeric(as.character(YR))) %>%
  ggplot(aes(x = Year, y = .value, colour = DEER, fill = DEER)) +
  stat_lineribbon(alpha = 1/4) +
  scale_fill_manual(values = two_cols,
                    aesthetics = c("colour","fill"),
                    name = "Deer Risk") +
  theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14)) +
  labs(y = "Forb cover")
```

### Fern cover

```{r fern model final}
fern_mod <- brm(bf(FERN_COVER ~ -1 + YR*mo(DEER) + YDAY + (1|SITE/PLOT),
                   hu ~ -1 + YR*mo(DEER) + YDAY + (1|SITE/PLOT)),
                data = BRLEAF_SUMMARY, family = "hurdle_gamma",
                prior = mod_pr, cores = 4, warmup = 2000, iter = 6000, thin = 4,
                file = paste0(model_loc, "FERN_COVER"),
                control = list(adapt_delta = 0.9))
par_summary(fern_mod)
```

```{r fern model plot diag, fig.dim = c(7,8)}
mcmc_plot(fern_mod, type = "rank_overlay")
pp_check(fern_mod, type = "ecdf_overlay", ndraws = 20) +
  scale_x_continuous(limits = c(0,250))
resid_bysite(fern_mod)
```

```{r emmeans fern pairwise comparison table}
fern_emm <- emmeans(fern_mod, ~ DEER | YR, epred = TRUE)
pairs(fern_emm) %>%
  knitr::kable(digits = 3, caption = "Fern cover")
```

```{r emmeans fern plot}
fern_emm %>%
  gather_emmeans_draws() %>%
  mutate(Year = as.numeric(as.character(YR))) %>%
  ggplot(aes(x = Year, y = .value, colour = DEER, fill = DEER)) +
  stat_lineribbon(alpha = 1/4) +
  scale_fill_manual(values = two_cols,
                    aesthetics = c("colour","fill"),
                    name = "Deer Risk") +
  theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14)) +
  labs(y = "Fern cover")
```

### Grass cover

```{r grass model final}
grass_mod <- brm(bf(GRASS_COVER ~ -1 + YR*mo(DEER) + YDAY + (1|SITE/PLOT),
                    hu ~ -1 + YR*mo(DEER) + YDAY + (1|SITE/PLOT)),
                 data = BRLEAF_SUMMARY, family = "hurdle_gamma",
                 prior = mod_pr, cores = 4, warmup = 2000, iter = 6000, thin = 4,
                 file = paste0(model_loc, "GRASS_COVER"))
par_summary(grass_mod)
```

```{r grass model plot diag, fig.dim = c(7,8)}
mcmc_plot(grass_mod, type = "rank_overlay")
pp_check(grass_mod, type = "ecdf_overlay", ndraws = 20) +
  scale_x_continuous(limits = c(0,250))
resid_bysite(grass_mod)
```

```{r emmeans grass pairwise comparison table}
grass_emm <- emmeans(grass_mod, ~ DEER | YR, epred = TRUE)
pairs(grass_emm) %>%
  knitr::kable(digits = 3, caption = "Grass cover")
```

```{r emmeans grass plot}
grass_emm %>%
  gather_emmeans_draws() %>%
  mutate(Year = as.numeric(as.character(YR))) %>%
  ggplot(aes(x = Year, y = .value, colour = DEER, fill = DEER)) +
  stat_lineribbon(alpha = 1/4) +
  scale_fill_manual(values = two_cols,
                    aesthetics = c("colour","fill"),
                    name = "Deer Risk") +
  theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14)) +
  labs(y = "Grass cover")
```

```{r emmeans grass pairwise comparison table mu}
grass_emm <- emmeans(grass_mod, ~ DEER | YR, dpar = "mu")
pairs(grass_emm, type = "response") %>%
  knitr::kable(digits = 3, caption = "Grass cover on log-scale")
```

```{r emmeans grass pairwise comparison table hu}
grass_emm <- emmeans(grass_mod, ~ DEER | YR, dpar = "hu")
pairs(grass_emm, type = "response") %>%
  knitr::kable(digits = 3, caption = "Grass cover on log-scale")
```


### Wood cover

```{r wood model final}
wood_mod <- brm(bf(WOOD_COVER ~ -1 + YR*mo(DEER) + YDAY + (1|SITE/PLOT),
                   hu ~ -1 + YR*mo(DEER) + YDAY + (1|SITE/PLOT)),
                data = BRLEAF_SUMMARY, family = "hurdle_gamma",
                prior = mod_pr, cores = 4, warmup = 2000, iter = 6000, thin = 4,
                file = paste0(model_loc, "WOOD_COVER"))
par_summary(wood_mod)
```

```{r wood model plot diag, fig.dim = c(7,8)}
mcmc_plot(wood_mod, type = "rank_overlay")
pp_check(wood_mod, type = "ecdf_overlay", ndraws = 20) +
  scale_x_continuous(limits = c(0,250))
resid_bysite(wood_mod)
```

```{r emmeans wood pairwise comparison table}
wood_emm <- emmeans(wood_mod, ~ DEER | YR, epred = TRUE)
pairs(wood_emm) %>%
  knitr::kable(digits = 3, caption = "Wood cover")
```

```{r emmeans wood plot}
wood_emm %>%
  gather_emmeans_draws() %>%
  mutate(Year = as.numeric(as.character(YR))) %>%
  ggplot(aes(x = Year, y = .value, colour = DEER, fill = DEER)) +
  stat_lineribbon(alpha = 1/4) +
  scale_fill_manual(values = two_cols,
                    aesthetics = c("colour","fill"),
                    name = "Deer Risk") +
  theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14)) +
  labs(y = "Wood cover")
```


## Bramble cover

```{r rubus model final}
rubus_mod <- brm(bf(RUBUS ~ -1 + YR*mo(DEER) + YDAY + (1|SITE/PLOT),
                    hu ~ -1 + YR*mo(DEER) + YDAY + (1|SITE/PLOT)),
                 data = BRLEAF_SUMMARY, family = "hurdle_gamma",
                 prior = mod_pr, cores = 4, warmup = 2000, iter = 6000, thin = 4,
                 file = paste0(model_loc, "RUBUS"),
                 control = list(adapt_delta = 0.9))
par_summary(rubus_mod)
```

```{r rubus model plot diag, fig.dim = c(7,8)}
mcmc_plot(rubus_mod, type = "rank_overlay")
pp_check(rubus_mod, type = "ecdf_overlay", ndraws = 20) +
  scale_x_continuous(limits = c(0,150))
resid_bysite(rubus_mod)
```

```{r emmeans rubus pairwise comparison table year}
rubus_emm <- emmeans(rubus_mod, ~ YR | DEER, epred = TRUE)
pairs(rubus_emm) %>%
  knitr::kable(digits = 3)
```

```{r emmeans rubus pairwise comparison table}
rubus_emm <- emmeans(rubus_mod, ~ DEER | YR, epred = TRUE)
pairs(rubus_emm) %>%
  knitr::kable(digits = 3)
```

```{r emmeans rubus plot}
rubus_emm %>%
  gather_emmeans_draws() %>%
  mutate(Year = as.numeric(as.character(YR))) %>%
  ggplot(aes(x = Year, y = .value, colour = DEER, fill = DEER)) +
  stat_lineribbon(alpha = 1/4) +
  scale_fill_manual(values = two_cols,
                    aesthetics = c("colour","fill"),
                    name = "Deer Risk") +
  theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14)) +
  labs(y = "Bramble cover")
```

# Binomial models 

## AWI proportion

```{r awi prop model final}
mod_pr <- prior(normal(0,1), class = "b", coef = "YR1971") +
  prior(normal(0,1), class = "b", coef = "YR2001") +
  prior(normal(0,1), class = "b", coef = "YR2022") +
  prior(normal(0,1), class = "b", coef = "YDAY") +
  prior(normal(0,1), class = "b", coef = "moDEER") +
  prior(normal(0,1), class = "b", coef = "moDEER:YR2001") +
  prior(normal(0,1), class = "b", coef = "moDEER:YR2022") +
  prior(dirichlet(1), class = "simo", coef = "moDEER1") +
  prior(dirichlet(1), class = "simo", coef = "moDEER:YR20011") +
  prior(dirichlet(1), class = "simo", coef = "moDEER:YR20221") +
  prior(student_t(5, 0, 1), class = "sd")
awi_mod2 <- brm(AWI_RICH | trials(SPECIES_RICH) ~ -1 + YR*mo(DEER) + YDAY + (1|SITE/PLOT) + (1|AWI_region),
                data = BRLEAF_SUMMARY, family = binomial(),
                prior = mod_pr, 
                cores = 4, warmup = 2000, iter = 6000, thin = 4,
                control = list(adapt_delta = 0.99),
                file = paste0(model_loc, "AWI_RICH_BIN"))
par_summary(awi_mod2)
```

```{r awi prop model plots}
mcmc_plot(awi_mod2, type = "rank_overlay")
pp_check(awi_mod2, type = "ecdf_overlay")
resid_bysite(awi_mod2)
```

```{r emmeans awi prop pairwise comparison table}
awi_emm <- emmeans(awi_mod2, ~ DEER | YR)
pairs(awi_emm, type = "response") %>%
  knitr::kable(digits = 3)
```

```{r emmeans awi prop plot}
awi_emm %>%
  gather_emmeans_draws() %>%
  mutate(Year = as.numeric(as.character(YR)),
         .value = inv_logit_scaled(.value)) %>%
  ggplot(aes(x = Year, y = .value, colour = DEER, fill = DEER)) +
  stat_lineribbon(alpha = 1/4) +
  scale_fill_manual(values = two_cols,
                    aesthetics = c("colour","fill"),
                    name = "Deer Risk") +
  theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14)) +
  labs(y = "AWI proportion")
```


# Binary models - regeneration

```{r regen data prep}
REGEN <- read.csv("Outputs/REGEN.csv")
REGEN <- REGEN %>%
  inner_join(select(BRLEAF_SUMMARY, SITE_NO, PLOT_NO, YEAR, YDAY, SITE, PLOT, DEER,YR)) 
```

```{r bin model sample prior}
mod_pr <- prior(normal(0,1), class = "b", coef = "YR1971") +
  prior(normal(0,1), class = "b", coef = "YR2001") +
  prior(normal(0,1), class = "b", coef = "YR2022") +
  prior(normal(0,1), class = "b") +
  prior(student_t(5, 0, 1), class = "sd")
```

```{r bin check prior predictions, cache = TRUE}
test_mod <- brm(REGEN ~ -1 + YR*mo(DEER) + (1|SITE/PLOT),
                data = REGEN, family = "bernoulli",
                prior = mod_pr, cores = 4, sample_prior = "only")
plot(test_mod)
pp_check(test_mod)
pp_check(test_mod, "ecdf_overlay", ndraws = 20)
```



## Ash regeneration

```{r ash model final}
ash_mod <- brm(ASH_REGEN ~ -1 + YR*mo(DEER) + (1|SITE/PLOT),
               data = REGEN, family = "bernoulli",
               prior = mod_pr, cores = 4, warmup = 2000, iter = 6000, thin = 4,
               file = paste0(model_loc, "ASH_REGEN"))
par_summary(ash_mod)
```

```{r ash model plots}
mcmc_plot(ash_mod, type = "rank_overlay")
pp_check(ash_mod, type = "stat")
resid_bysite(ash_mod)
```

```{r emmeans ash pairwise comparison table}
ash_emm <- emmeans(ash_mod, ~ DEER | YR)
pairs(ash_emm, type = "response") %>%
  knitr::kable(digits = 3)
```

```{r emmeans ash plot}
ash_emm %>%
  gather_emmeans_draws() %>%
  mutate(Year = as.numeric(as.character(YR)),
         .value = inv_logit_scaled(.value)) %>%
  ggplot(aes(x = Year, y = .value, colour = DEER, fill = DEER)) +
  stat_lineribbon(alpha = 1/4) +
  scale_fill_manual(values = two_cols,
                    aesthetics = c("colour","fill"),
                    name = "Deer Risk") +
  theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14)) +
  labs(y = "Ash regeneration")
```

## Total regeneration

```{r regen model final}
regen_mod <- brm(REGEN ~ -1 + YR*mo(DEER) + (1|SITE/PLOT),
                 data = REGEN, family = "bernoulli",
                 prior = mod_pr, cores = 4, warmup = 2000, iter = 6000, thin = 4,
                 file = paste0(model_loc, "REGEN"))
par_summary(regen_mod)
```

```{r regen model plots}
mcmc_plot(regen_mod, type = "rank_overlay")
pp_check(regen_mod, type = "stat") 
resid_bysite(regen_mod)
```

```{r emmeans regen pairwise comparison table}
regen_emm <- emmeans(regen_mod, ~ DEER | YR)
pairs(regen_emm, type = "response") %>%
  knitr::kable(digits = 3)
```

```{r emmeans regen plot}
regen_emm %>%
  gather_emmeans_draws() %>%
  mutate(Year = as.numeric(as.character(YR)),
         .value = inv_logit_scaled(.value)) %>%
  ggplot(aes(x = Year, y = .value, colour = DEER, fill = DEER)) +
  stat_lineribbon(alpha = 1/4) +
  scale_fill_manual(values = two_cols,
                    aesthetics = c("colour","fill"),
                    name = "Deer Risk") +
  theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14)) +
  labs(y = "Total regeneration")
```

```{r emmeans by year regen pairwise comparison table}
regen_emm <- emmeans(regen_mod, ~ YR | DEER)
pairs(regen_emm, type = "response") %>%
  knitr::kable(digits = 3)
```


# Site level richness

```{r site level data prep}
BRLEAF_SUMM_SITE <- read.csv("Outputs/Site level richness.csv")
BRLEAF_SUMM_SITE <- BRLEAF_SUMM_SITE %>%
  inner_join(deerrisk, by = c("SITE_NO" = "SITE_NUMBER")) %>%
  mutate(YR = c("1971","2001","2022")[YEAR],
         YDAY = (DAYOFYEAR - 200)/30,
         SITE = as.character(SITE_NO),
         DEER = factor(DEER, c("Low","Moderate","High"),
                       ordered = TRUE))
```

```{r count model sample prior site level}
mod_pr <- prior(normal(2,1), class = "b", coef = "YR1971") +
  prior(normal(2,1), class = "b", coef = "YR2001") +
  prior(normal(2,1), class = "b", coef = "YR2022") +
  prior(normal(0,1), class = "b", coef = "YDAY") +
  prior(normal(0,1), class = "b", coef = "moDEER") +
  prior(normal(0,1), class = "b", coef = "moDEER:YR2001") +
  prior(normal(0,1), class = "b", coef = "moDEER:YR2022") +
  prior(dirichlet(1), class = "simo", coef = "moDEER1") +
  prior(dirichlet(1), class = "simo", coef = "moDEER:YR20011") +
  prior(dirichlet(1), class = "simo", coef = "moDEER:YR20221") +
  prior(student_t(5, 0, 1), class = "sd")
```

```{r site rich mod run}
site_rich_mod <- brm(SPECIES_RICH ~ -1 + YR*mo(DEER) + YDAY + (1|SITE),
                     data = BRLEAF_SUMM_SITE, family = "negbinomial",
                     prior = mod_pr, cores = 4, warmup = 2000, iter = 6000, thin = 4,
                     file = paste0(model_loc, "SITERICH_NB"))
par_summary(site_rich_mod)
```

```{r site rich mod plots}
mcmc_plot(site_rich_mod, type = "rank_overlay")
pp_check(site_rich_mod, type = "ecdf_overlay", ndraws = 20)
resid_bysite(site_rich_mod)
```

```{r emmeans site richness pairwise comparison table}
siterich_emm <- emmeans(site_rich_mod, ~ DEER | YR)
pairs(siterich_emm, type = "response") %>%
  knitr::kable(digits = 3, caption = "Site level richness on response-scale")
```

```{r emmeans site richness plot}
siterich_emm %>%
  gather_emmeans_draws() %>%
  mutate(Year = as.numeric(as.character(YR)),
         .value = exp(.value)) %>%
  ggplot(aes(x = Year, y = .value, colour = DEER, fill = DEER)) +
  stat_lineribbon(alpha = 1/4) +
  scale_fill_manual(values = two_cols,
                    aesthetics = c("colour","fill"),
                    name = "Deer Risk") +
  theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14)) +
  labs(y = "Site-level Species Richness")
```


# Site level AWI proportion

```{r site level awi data prep}
GRFLORA_SITE <- read.csv("Outputs/Site level richness with AWI.csv")
BRLEAF_SUMMARY_SITE <- GRFLORA_SITE %>%
  mutate(YR = c("1971","2001","2022")[YEAR],
         YDAY = (DAYOFYEAR - 200)/30,
         SITE = as.character(SITE_NO)) %>%
  left_join(select(BRLEAF_SUMMARY, SITE_NO, AWI_region) %>% distinct()) %>%
  inner_join(deerrisk) %>%
  mutate(DEER = factor(DEER, c("Low","Moderate","High"),
                       ordered = TRUE))
```

```{r binomial model sample prior site level}
mod_pr <- prior(normal(0,1), class = "b") +
  prior(student_t(5, 0, 1), class = "sd")
```

```{r awi site rich mod run}
awisite_mod <- brm(AWI_RICH | trials(SPECIES_RICH) ~ -1 + YR*mo(DEER) + YDAY + (1|SITE) + (1|AWI_region),
                   data = BRLEAF_SUMMARY_SITE, family = binomial(),
                   prior = mod_pr, cores = 4, warmup = 2000, iter = 6000, thin = 4,
                   control = list(adapt_delta = 0.95),
                   file = paste0(model_loc, "SITERICH_AWI_BIN"))
par_summary(awisite_mod)
```

```{r awi site rich mod plots}
mcmc_plot(awisite_mod, type = "rank_overlay")
pp_check(awisite_mod, type = "ecdf_overlay", ndraws = 20)
resid_bysite(awisite_mod)
```

```{r emmeans awi site richness pairwise comparison table}
awisite_emm <- emmeans(awisite_mod, ~ DEER | YR)
pairs(awisite_emm, type = "response") %>%
  knitr::kable(digits = 3, caption = "Site level AWI proportion on response-scale")
```

```{r emmeans awi site richness plot}
awisite_emm %>%
  gather_emmeans_draws() %>%
  mutate(Year = as.numeric(as.character(YR)),
         .value = exp(.value)) %>%
  ggplot(aes(x = Year, y = .value, colour = DEER, fill = DEER)) +
  stat_lineribbon(alpha = 1/4) +
  scale_fill_manual(values = two_cols,
                    aesthetics = c("colour","fill"),
                    name = "Deer Risk") +
  theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14)) +
  labs(y = "Site-level AWI proportion")
```
