---
title: "Multivariate Models"
author: "Anonymised"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, fig.dim = c(9,5))
```


```{r libraries}
library(dplyr)
library(brms)
library(emmeans)
library(tidybayes)
library(ggplot2)
theme_set(theme_classic())
library(patchwork)
teal <- "#0383a4"
two_cols <- c("#999999","#0072B2","#E69F00")
model_loc <- "Outputs/Models/Multivariate_Models/"
```

```{r, include = FALSE}
source("Helper_functions.R")
```




```{r deer risk data read, cache = TRUE}
deerrisk <- read.csv("Metadata/DEER_RISK.csv") %>%
  mutate(DEER = ifelse(DEER == "None", "Low", DEER)) %>%
  rename(SITE_NO = SITE_NUMBER)
regions <- read.csv("Outputs/regions.csv")
ash_plots <- read.csv("Outputs/ash_plots.csv")
ash_sites <- unique(ash_plots$SITE_NO)
```

```{r data read}
BRLEAF_SUMMARY <- read.csv("Outputs/Broadleaf_summary_metrics.csv")
BRLEAF_SUMMARY <- BRLEAF_SUMMARY %>%
  inner_join(deerrisk, by = "SITE_NO") %>%
  left_join(ash_plots) %>%
  filter(!is.na(ASHDIEBACK) & SITE_NO %in% ash_sites) %>%
  mutate(ASHPRESENT = tidyr::replace_na(ASHPRESENT, FALSE)) %>%
  mutate(ASHDIEBACK = ifelse(ASHDIEBACK == "Present", "Present",
                             ifelse(ASHPRESENT & ASHDIEBACK == "Not present",
                                    "NotPresent", "NoAsh")),
         ASH = ifelse(ASHPRESENT | ASHDIEBACK == "Present", "Ash", "NoAsh"),
         YR = c("1971","2001","2022")[YEAR],
         YDAY = (DAYOFYEAR - 200)/30,
         SITE = as.character(SITE_NO),
         PLOT = paste(SITE_NO, PLOT_NO, sep = "_"),
         DEER = factor(DEER, c("Low","Moderate","High"),
                       ordered = TRUE))
Climate_data <- read.csv("Outputs/Climate_data.csv") %>%
  tidyr::pivot_longer(Winter_rainfall_2022:Summer_tasmax_2001,
                      names_to = c("Season","Variable","YR"),
                      names_sep = "_") %>%
  mutate(Clim_var = paste(Season, Variable, sep = "_")) %>%
  select(SITE_NO, PLOT_NO, YR, Clim_var, value) %>% distinct() %>%
  tidyr::pivot_wider(names_from = Clim_var, values_from = value)
BRLEAF_SUMMARY <- inner_join(BRLEAF_SUMMARY, Climate_data) %>%
  mutate(Summer_rainfall = (Summer_rainfall - 250)/100,
         Winter_rainfall = (Winter_rainfall - 300)/100,
         Summer_tasmax = (Summer_tasmax - 19),
         Winter_tasmin = (Winter_tasmin - 1))
REGEN <- read.csv("Outputs/REGEN.csv") %>%
  select(SITE_NO, PLOT_NO, YEAR, ends_with("REGEN")) %>% distinct()
BRLEAF_SUMMARY <- inner_join(BRLEAF_SUMMARY, REGEN)
```

# Multivariate Model

## Full model


```{r model structure}
model_structure <-
  bf(SPECIES_RICH ~ YR*ASHDIEBACK*mo(DEER) + YDAY + FORB_COVER + GRASS_COVER + RUBUS +
       Winter_rainfall + Summer_rainfall + Winter_tasmin + Summer_tasmax + 
       (1|SITE/PLOT), family = "negbinomial") +
  bf(REGEN ~ YR*ASHDIEBACK*mo(DEER) + YDAY + FORB_COVER + GRASS_COVER + RUBUS +
       Winter_rainfall + Summer_rainfall + Winter_tasmin + Summer_tasmax + 
       (1|SITE/PLOT), family = "bernoulli") +
  bf(FORB_COVER ~ YR*ASHDIEBACK + YDAY + RUBUS + 
       Winter_rainfall + Summer_rainfall + Winter_tasmin + Summer_tasmax + 
       (1|SITE/PLOT),
     hu ~ YR*ASHDIEBACK + YDAY + RUBUS + 
       Winter_rainfall + Summer_rainfall + Winter_tasmin + Summer_tasmax + 
       (1|SITE/PLOT), family = "hurdle_gamma") +
  bf(GRASS_COVER ~ YR*mo(DEER) + YDAY + RUBUS + 
       Winter_rainfall + Summer_rainfall + Winter_tasmin + Summer_tasmax + 
       (1|SITE/PLOT),
     hu ~ YR*mo(DEER) + YDAY + RUBUS + 
       Winter_rainfall + Summer_rainfall + Winter_tasmin + Summer_tasmax + 
       (1|SITE/PLOT), family = "hurdle_gamma") +
  bf(RUBUS ~ YR*mo(DEER) + YDAY + 
       Winter_rainfall + Summer_rainfall + Winter_tasmin + Summer_tasmax + 
       (1|SITE/PLOT), 
     hu ~ YR*mo(DEER) + YDAY + 
       Winter_rainfall + Summer_rainfall + Winter_tasmin + Summer_tasmax + 
       (1|SITE/PLOT), family = "hurdle_gamma") +
  set_rescor(FALSE)
```

```{r specify priors}
mod_pr_nodeer <- 
  # forb priors
  prior(student_t(4,2,1), class = "Intercept", resp = "FORBCOVER") +
  prior(normal(0,1), class = "b", resp = "FORBCOVER") +
  prior(student_t(5, 0, 1), class = "sd", resp = "FORBCOVER") +
  prior(gamma(0.1,0.1), class = "shape", resp = "FORBCOVER") +
  prior(student_t(4,-2,1), class = "Intercept", dpar = "hu", resp = "FORBCOVER") +
  prior(normal(0,1), class = "b", dpar = "hu", resp = "FORBCOVER") +
  prior(student_t(2, 0, 1), class = "sd", dpar = "hu", resp = "FORBCOVER") +
  # grass priors
  prior(student_t(4,2,1), class = "Intercept", resp = "GRASSCOVER") +
  prior(normal(0,1), class = "b", resp = "GRASSCOVER") +
  prior(student_t(5, 0, 1), class = "sd", resp = "GRASSCOVER") +
  prior(dirichlet(1), class = "simo", resp = "GRASSCOVER", coef = "moDEER1") +
  prior(gamma(0.1,0.1), class = "shape", resp = "GRASSCOVER") +
  prior(student_t(4,-2,1), class = "Intercept",  dpar = "hu", resp = "GRASSCOVER") +
  prior(normal(0,1), class = "b", dpar = "hu", resp = "GRASSCOVER") +
  prior(student_t(2, 0, 1), class = "sd", dpar = "hu", resp = "GRASSCOVER") +
  prior(dirichlet(1), class = "simo", dpar = "hu", resp = "GRASSCOVER", coef = "moDEER1") +
  # rubus priors
  prior(student_t(4,2,1), class = "Intercept", resp = "RUBUS") +
  prior(normal(0,1), class = "b", resp = "RUBUS") +
  prior(student_t(5, 0, 1), class = "sd", resp = "RUBUS") +
  prior(dirichlet(1), class = "simo", resp = "RUBUS", coef = "moDEER1") +
  prior(gamma(0.1,0.1), class = "shape", resp = "RUBUS") +
  prior(student_t(4,-1,1), class = "Intercept", dpar = "hu", resp = "RUBUS") +
  prior(normal(0,1), class = "b", dpar = "hu", resp = "RUBUS") +
  prior(student_t(2, 0, 1), class = "sd", dpar = "hu", resp = "RUBUS") +
  prior(dirichlet(1), class = "simo", dpar = "hu", resp = "RUBUS", coef = "moDEER1") +
  # species richness priors
  prior(student_t(4,3,1), class = "Intercept", resp = "SPECIESRICH") +
  prior(normal(0,1), class = "b", resp = "SPECIESRICH") +
  prior(student_t(5, 0, 1), class = "sd", resp = "SPECIESRICH") +
  prior(gamma(0.1,0.1), class = "shape", resp = "SPECIESRICH") +
  # regeneration priors
  prior(student_t(4,0,1), class = "Intercept", resp = "REGEN") +
  prior(normal(0,1), class = "b", resp = "REGEN") +
  prior(student_t(5, 0, 1), class = "sd", resp = "REGEN") 
mod_pr_deer <- mod_pr_nodeer +
  prior(dirichlet(1), class = "simo", resp = "SPECIESRICH", coef = "moDEER1") +
  prior(dirichlet(1), class = "simo", resp = "REGEN", coef = "moDEER1")
```


```{r run full model}
sem_allclim <- brm(model_structure, data = BRLEAF_SUMMARY, prior = mod_pr_deer,
                   iter = 6000, warmup = 2000, thin = 4, cores = 4,
                   file = paste0(model_loc, "SEM_AllClim"))
par_summary(sem_allclim)
```

```{r prior retrodictive check}
pp_check(sem_allclim, resp = "FORBCOVER", ndraws = 20, type = "ecdf_overlay") + 
  scale_x_continuous(limits = c(0,300))
pp_check(sem_allclim, resp = "GRASSCOVER", ndraws = 20, type = "ecdf_overlay") + 
  scale_x_continuous(limits = c(0,300))
pp_check(sem_allclim, resp = "RUBUS", ndraws = 20, type = "ecdf_overlay") +
  scale_x_continuous(limits = c(0,200))
pp_check(sem_allclim, resp = "SPECIESRICH", ndraws = 20, type = "ecdf_overlay") +
  scale_x_continuous(limits = c(0,100))
pp_check(sem_allclim, resp = "REGEN", ndraws = 20, type = "ecdf_overlay") 
```

```{r}
resid_bysite_mult(sem_allclim)
```



```{r full model kfold}
set.seed(1)
folds <- loo::kfold_split_grouped(K = 10, x = BRLEAF_SUMMARY$SITE)
sem_allclim <- add_criterion(sem_allclim, "kfold", folds = folds,
                             file = paste0(model_loc, "SEM_FullModel"))
```

## Run climate models

### Model with only winter rainfall and winter tasmin

```{r all winter model structure}
model_structure <-
  bf(SPECIES_RICH ~ YR*ASHDIEBACK*mo(DEER) + YDAY + FORB_COVER + GRASS_COVER + RUBUS +
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), family = "negbinomial") +
  bf(REGEN ~ YR*ASHDIEBACK*mo(DEER) + YDAY + FORB_COVER + GRASS_COVER + RUBUS +
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), family = "bernoulli") +
  bf(FORB_COVER ~ YR*ASHDIEBACK + YDAY + RUBUS + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT),
     hu ~ YR*ASHDIEBACK + YDAY + RUBUS + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), family = "hurdle_gamma") +
  bf(GRASS_COVER ~ YR*mo(DEER) + YDAY + RUBUS + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT),
     hu ~ YR*mo(DEER) + YDAY + RUBUS + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), family = "hurdle_gamma") +
  bf(RUBUS ~ YR*mo(DEER) + YDAY + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), 
     hu ~ YR*mo(DEER) + YDAY + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), family = "hurdle_gamma") +
  set_rescor(FALSE)
```


```{r run all winter model}
sem_winclim <- brm(model_structure, data = BRLEAF_SUMMARY, prior = mod_pr_deer,
                   iter = 6000, warmup = 2000, thin = 4, cores = 4,
                   file = paste0(model_loc, "SEM_WinterClim"))
par_summary(sem_winclim)
```

```{r all winter model kfold}
sem_winclim <- add_criterion(sem_winclim, "kfold", folds = folds,
                             file = paste0(model_loc, "SEM_WinterClim"))
```

### Model with only summer rainfall and summer tasmax

```{r all summer model structure}
model_structure <-
  bf(SPECIES_RICH ~ YR*ASHDIEBACK*mo(DEER) + YDAY + FORB_COVER + GRASS_COVER + RUBUS + 
       Summer_rainfall + Summer_tasmax + 
       (1|SITE/PLOT), family = "negbinomial") +
  bf(REGEN ~ YR*ASHDIEBACK*mo(DEER) + YDAY + FORB_COVER + GRASS_COVER + RUBUS +
       Summer_rainfall + Summer_tasmax + 
       (1|SITE/PLOT), family = "bernoulli") +
  bf(FORB_COVER ~ YR*ASHDIEBACK + YDAY + RUBUS + 
       Summer_rainfall + Summer_tasmax + 
       (1|SITE/PLOT),
     hu ~ YR*ASHDIEBACK + YDAY + RUBUS +
       Summer_rainfall + Summer_tasmax + 
       (1|SITE/PLOT), family = "hurdle_gamma") +
  bf(GRASS_COVER ~ YR*mo(DEER) + YDAY + RUBUS + 
       Summer_rainfall + Summer_tasmax + 
       (1|SITE/PLOT),
     hu ~ YR*mo(DEER) + YDAY + RUBUS + 
       Summer_rainfall + Summer_tasmax + 
       (1|SITE/PLOT), family = "hurdle_gamma") +
  bf(RUBUS ~ YR*mo(DEER) + YDAY + 
       Summer_rainfall + Summer_tasmax + 
       (1|SITE/PLOT), 
     hu ~ YR*mo(DEER) + YDAY + 
       Summer_rainfall + Summer_tasmax + 
       (1|SITE/PLOT), family = "hurdle_gamma") +
  set_rescor(FALSE)
```

```{r run all summer model}
sem_sumclim <- brm(model_structure, data = BRLEAF_SUMMARY, prior = mod_pr_deer,
                   iter = 6000, warmup = 2000, thin = 4, cores = 4,
                   file = paste0(model_loc, "SEM_SummerClim"))
par_summary(sem_sumclim)
```

```{r all summer model kfold}
sem_sumclim <- add_criterion(sem_sumclim, "kfold", folds = folds,
                             file = paste0(model_loc, "SEM_SummerClim"))
```


### Model with no climate variables

```{r model structure no climate}
model_structure <-
  bf(SPECIES_RICH ~ -1 + YR*ASHDIEBACK*mo(DEER) + YDAY + FORB_COVER + GRASS_COVER + RUBUS + 
       (1|SITE/PLOT), family = "negbinomial") +
  bf(REGEN ~ -1 + YR*ASHDIEBACK*mo(DEER) + YDAY + FORB_COVER + GRASS_COVER + RUBUS +
       (1|SITE/PLOT), family = "bernoulli") +
  bf(FORB_COVER ~ -1 + YR*ASHDIEBACK + YDAY + RUBUS + 
       (1|SITE/PLOT),
     hu ~ -1 + YR*ASHDIEBACK + YDAY + RUBUS + 
       (1|SITE/PLOT), family = "hurdle_gamma") +
  bf(GRASS_COVER ~ -1 + YR*mo(DEER) + YDAY + RUBUS +
       (1|SITE/PLOT),
     hu ~ -1 + YR*mo(DEER) + YDAY + RUBUS + 
       (1|SITE/PLOT), family = "hurdle_gamma") +
  bf(RUBUS ~ -1 + YR*mo(DEER) + YDAY + 
       (1|SITE/PLOT), 
     hu ~ -1 + YR*mo(DEER) + YDAY + 
       (1|SITE/PLOT), family = "hurdle_gamma") +
  set_rescor(FALSE)
```

```{r run no climate model}
sem_noclim <- brm(model_structure, data = BRLEAF_SUMMARY, prior = mod_pr_deer,
                  iter = 6000, warmup = 2000, thin = 4, cores = 4,
                  file = paste0(model_loc, "SEM_NoClim"))
par_summary(sem_noclim)
```

```{r no climate model kfold}
sem_noclim <- add_criterion(sem_noclim, "kfold", folds = folds,
                            file = paste0(model_loc, "SEM_NoClim"))
```



## Compare models with kfold

```{r kfold compare}
loo_compare(sem_allclim, sem_winclim, sem_sumclim, sem_noclim,
            criterion = "kfold") %>%
  knitr::kable(digits = 3)
```

Model with winter climate only performs the best out of all model options, so all further model testing will be done with winter climate only.



## Test impact of removing direct deer and ash dieback links to species richness and tree regeneration

### Winter Model with no deer 

```{r winter no deer model structure}
model_structure <-
  bf(SPECIES_RICH ~ YR*ASHDIEBACK + YDAY + FORB_COVER + GRASS_COVER + RUBUS + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), family = "negbinomial") +
  bf(REGEN ~ YR*ASHDIEBACK + YDAY + FORB_COVER + GRASS_COVER + RUBUS +
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), family = "bernoulli") +
  bf(FORB_COVER ~ YR*ASHDIEBACK + YDAY + RUBUS +
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT),
     hu ~ YR*ASHDIEBACK + YDAY + RUBUS +
       Winter_rainfall + Winter_tasmin +
       (1|SITE/PLOT), family = "hurdle_gamma") +
  bf(GRASS_COVER ~ YR*mo(DEER) + YDAY + RUBUS + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT),
     hu ~ YR*mo(DEER) + YDAY + RUBUS + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), family = "hurdle_gamma") +
  bf(RUBUS ~ YR*mo(DEER) + YDAY + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), 
     hu ~ YR*mo(DEER) + YDAY + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), family = "hurdle_gamma") +
  set_rescor(FALSE)
```


```{r run no deer winter model}
sem_win_nodeer <- brm(model_structure, data = BRLEAF_SUMMARY, prior = mod_pr_nodeer,
                      iter = 6000, warmup = 2000, thin = 4, cores = 4,
                      file = paste0(model_loc, "SEM_WinterClim_DeerMed"))
par_summary(sem_win_nodeer)
```


```{r no deer winter model kfold}
sem_win_nodeer <- add_criterion(sem_win_nodeer, "kfold", folds = folds,
                                file = paste0(model_loc, "SEM_WinterClim_DeerMed"))
```

### Winter Model with no dieback


```{r no dieback winter model structure}
model_structure <-
  bf(SPECIES_RICH ~ YR*mo(DEER)*ASH + YDAY + FORB_COVER + GRASS_COVER + RUBUS + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), family = "negbinomial") +
  bf(REGEN ~ YR*mo(DEER)*ASH + YDAY + FORB_COVER + GRASS_COVER + RUBUS +
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), family = "bernoulli") +
  bf(FORB_COVER ~ YR*ASHDIEBACK + YDAY + RUBUS +
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT),
     hu ~ YR*ASHDIEBACK + YDAY + RUBUS +
       Winter_rainfall + Winter_tasmin +
       (1|SITE/PLOT), family = "hurdle_gamma") +
  bf(GRASS_COVER ~ YR*ASHDIEBACK + YDAY + RUBUS + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT),
     hu ~ YR*ASHDIEBACK + YDAY + RUBUS + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), family = "hurdle_gamma") +
  bf(RUBUS ~ YR*mo(DEER) + YDAY + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), 
     hu ~ YR*mo(DEER) + YDAY + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), family = "hurdle_gamma") +
  set_rescor(FALSE)
```


```{r run no dieback winter model}
sem_win_noDB <- brm(model_structure, data = BRLEAF_SUMMARY, prior = mod_pr_deer,
                    iter = 6000, warmup = 2000, thin = 4, cores = 4,
                    file = paste0(model_loc, "SEM_WinterClim_AshDBMed"))
par_summary(sem_win_noDB)
```

```{r no dieback winter model kfold}
sem_win_noDB <- add_criterion(sem_win_noDB, "kfold", folds = folds,
                              file = paste0(model_loc, "SEM_WinterClim_AshDBMed"))
```


### Winter Model with no dieback or deer

```{r no deer or dieback winter model structure}
model_structure <-
  bf(SPECIES_RICH ~ YR*ASH + YDAY + FORB_COVER + RUBUS + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), family = "negbinomial") +
  bf(REGEN ~ YR*ASH + YDAY + FORB_COVER + GRASS_COVER + RUBUS +
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), family = "bernoulli") +
  bf(FORB_COVER ~ YR*ASHDIEBACK + YDAY + RUBUS +
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT),
     hu ~ YR*ASHDIEBACK + YDAY + RUBUS +
       Winter_rainfall + Winter_tasmin +
       (1|SITE/PLOT), family = "hurdle_gamma") +
  bf(GRASS_COVER ~ YR*mo(DEER) + YDAY + RUBUS + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT),
     hu ~ YR*mo(DEER) + YDAY + RUBUS + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), family = "hurdle_gamma") +
  bf(RUBUS ~ YR*mo(DEER) + YDAY + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), 
     hu ~ YR*mo(DEER) + YDAY + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), family = "hurdle_gamma") +
  set_rescor(FALSE)
```


```{r run no deer or dieback winter model}
sem_win_fullmed <- brm(model_structure, data = BRLEAF_SUMMARY, prior = mod_pr_nodeer,
                       iter = 6000, warmup = 2000, thin = 4, cores = 4,
                       file = paste0(model_loc, "SEM_WinterClim_FullMed"))
par_summary(sem_win_fullmed)
```


```{r no deer or dieback winter model kfold}
sem_win_fullmed <- add_criterion(sem_win_fullmed, "kfold", folds = folds,
                                 file = paste0(model_loc, "SEM_WinterClim_FullMed"))
```


### Winter Model no direct effects on species richness


```{r winter model sr med structure}
model_structure <-
  bf(SPECIES_RICH ~ YR*ASH + YDAY + FORB_COVER + GRASS_COVER + RUBUS +
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), family = "negbinomial") +
  bf(REGEN ~ YR*ASHDIEBACK*mo(DEER) + YDAY + FORB_COVER + GRASS_COVER + RUBUS +
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), family = "bernoulli") +
  bf(FORB_COVER ~ YR*ASHDIEBACK + YDAY + RUBUS + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT),
     hu ~ YR*ASHDIEBACK + YDAY + RUBUS + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), family = "hurdle_gamma") +
  bf(GRASS_COVER ~ YR*mo(DEER) + YDAY + RUBUS + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT),
     hu ~ YR*mo(DEER) + YDAY + RUBUS + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), family = "hurdle_gamma") +
  bf(RUBUS ~ YR*mo(DEER) + YDAY + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), 
     hu ~ YR*mo(DEER) + YDAY + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), family = "hurdle_gamma") +
  set_rescor(FALSE)
```


```{r run winter sr med model}
sem_win_srmed <- brm(model_structure, data = BRLEAF_SUMMARY, 
                     prior = mod_pr_nodeer +
                       prior(dirichlet(1), class = "simo", resp = "REGEN", coef = "moDEER1"),
                     iter = 6000, warmup = 2000, thin = 4, cores = 4,
                     file = paste0(model_loc, "SEM_WinterClim_SRMed"))
par_summary(sem_win_srmed)
```

```{r winter model kfold}
sem_win_srmed <- add_criterion(sem_win_srmed, "kfold", folds = folds,
                             file = paste0(model_loc, "SEM_WinterClim_SRMed"))
```



### Winter Model no direct effects on tree regeneration

```{r winter model regen med structure}
model_structure <-
  bf(SPECIES_RICH ~ YR*ASHDIEBACK*mo(DEER) + YDAY + FORB_COVER + GRASS_COVER + RUBUS +
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), family = "negbinomial") +
  bf(REGEN ~ YR*ASH + YDAY + FORB_COVER + GRASS_COVER + RUBUS +
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), family = "bernoulli") +
  bf(FORB_COVER ~ YR*ASHDIEBACK + YDAY + RUBUS + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT),
     hu ~ YR*ASHDIEBACK + YDAY + RUBUS + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), family = "hurdle_gamma") +
  bf(GRASS_COVER ~ YR*mo(DEER) + YDAY + RUBUS + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT),
     hu ~ YR*mo(DEER) + YDAY + RUBUS + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), family = "hurdle_gamma") +
  bf(RUBUS ~ YR*mo(DEER) + YDAY + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), 
     hu ~ YR*mo(DEER) + YDAY + 
       Winter_rainfall + Winter_tasmin + 
       (1|SITE/PLOT), family = "hurdle_gamma") +
  set_rescor(FALSE)
```


```{r run winter regen med model}
sem_win_regmed <- brm(model_structure, data = BRLEAF_SUMMARY, 
                     prior = mod_pr_nodeer +
                       prior(dirichlet(1), class = "simo", resp = "SPECIESRICH", coef = "moDEER1"),
                     iter = 6000, warmup = 2000, thin = 4, cores = 4,
                     file = paste0(model_loc, "SEM_WinterClim_RegenMed"))
par_summary(sem_win_regmed)
```

```{r winter regen med model kfold}
sem_win_regmed <- add_criterion(sem_win_regmed, "kfold", folds = folds,
                             file = paste0(model_loc, "SEM_WinterClim_RegenMed"))
```



## Model comparison with kfold

```{r}
loo_compare(sem_winclim, sem_win_nodeer, sem_win_noDB, sem_win_fullmed,
            sem_win_srmed, sem_win_regmed,
            criterion = "kfold") %>%
  knitr::kable(digits = 3)
```

Best model is winter-only climate with direct effect of deer but no direct effect of ash dieback upon species richness and tree regeneration. However, winter-only model with complete mediation of deer and ash dieback was comparable in predictive performance and is a simpler model, so limiting to this model for summaries. 

# Model analysis

First, for the best model that includes direct deer effects upon species richness and tree regeneration graph those out, then do the rest of the work with the fully mediated model (effects are generally similar).

Deer effects:

```{r}
int_emmsr <- emmeans(sem_win_noDB, ~ DEER | YR, resp = "SPECIESRICH", epred = TRUE)
pairs(int_emmsr) %>%
  knitr::kable(digits = 3)
```

```{r}
(i1 <- int_emmsr %>%
  gather_emmeans_draws() %>%
  mutate(Year = as.numeric(as.character(YR))) %>%
  ggplot(aes(x = Year, y = .value, colour = DEER, fill = DEER)) +
  stat_pointinterval(position = position_dodge(width = 8)) +
  scale_fill_manual(values = two_cols,
                    aesthetics = c("colour","fill"),
                    name = "Deer Risk") +
  labs(y = "Ground Flora Richness") +
  theme(axis.text = element_text(size = 12)))
```

```{r}
int_emmreg <- emmeans(sem_win_noDB, ~ DEER | YR, resp = "REGEN", epred = TRUE)
pairs(int_emmreg) %>%
  knitr::kable(digits = 3)
```

```{r}
(i2 <- int_emmreg %>%
  gather_emmeans_draws() %>%
  mutate(Year = as.numeric(as.character(YR))) %>%
  ggplot(aes(x = Year, y = .value, colour = DEER, fill = DEER)) +
  stat_pointinterval(position = position_dodge(width = 8)) +
  scale_fill_manual(values = two_cols,
                    aesthetics = c("colour","fill"),
                    name = "Deer Risk") +
  labs(y = "Tree Regeneration") +
  theme(axis.text = element_text(size = 12)))
```



```{r}
intmodel <- sem_win_fullmed
```


Emmeans on ash dieback and deer effects:

```{r ash dieback effects on forbs}
int_emm1 <- emmeans(intmodel, ~ ASHDIEBACK | YR, resp = "FORBCOVER", epred = TRUE)
pairs(int_emm1) %>%
  knitr::kable(digits = 3)
```


```{r ash dieback effects on forbs plot, fig.dim = c(7,5)}
(i3 <- int_emm1 %>%
  gather_emmeans_draws() %>%
  mutate(Year = as.numeric(as.character(YR)),
         ASHDIEBACK = forcats::fct_recode(ASHDIEBACK, "Not Present" =  "NotPresent",
                                          "No Ash" = "NoAsh")) %>%
  ggplot(aes(x = Year, y = .value, colour = ASHDIEBACK, fill = ASHDIEBACK)) +
  stat_pointinterval(position = position_dodge(width = 8)) +
  scale_fill_manual(values = c("#000000","#009E73","#CC79A7"),
                    aesthetics = c("colour","fill"),
                    name = "Ash Dieback") +
  labs(y = "Forb cover (%)") +
  theme(axis.text = element_text(size = 12)))
```


```{r deer effects on grasses}
int_emm2 <- emmeans(intmodel, ~ DEER | YR, resp = "GRASSCOVER", epred = TRUE)
pairs(int_emm2) %>%
  knitr::kable(digits = 3)
```


```{r deer effects on grasses plot, fig.dim = c(7,5)}
(i4 <- int_emm2 %>%
  gather_emmeans_draws() %>%
  mutate(Year = as.numeric(as.character(YR))) %>%
  ggplot(aes(x = Year, y = .value, colour = DEER, fill = DEER)) +
  stat_pointinterval(position = position_dodge(width = 8)) +
  scale_fill_manual(values = two_cols,
                    aesthetics = c("colour","fill"),
                    name = "Deer Risk") +
  labs(y = "Grass cover (%)") +
  theme(axis.text = element_text(size = 12)))

```



```{r deer effects on bramble}
int_emm3 <- emmeans(intmodel, ~ DEER | YR, resp = "RUBUS", epred = TRUE)
pairs(int_emm3, type = "response") %>%
  knitr::kable(digits = 3)
```

```{r deer effects on rubus, fig.dim = c(7,5)}
(i5 <- int_emm3 %>%
  gather_emmeans_draws() %>%
  mutate(Year = as.numeric(as.character(YR))) %>%
  ggplot(aes(x = Year, y = .value, colour = DEER, fill = DEER)) +
  stat_pointinterval(position = position_dodge(width = 8)) +
  scale_fill_manual(values = two_cols,
                    aesthetics = c("colour","fill"),
                    name = "Deer Risk") +
  labs(y = "Bramble cover (%)") +
  theme(axis.text = element_text(size = 12)))

```




Effects of rubus on forb:

```{r bramble effect forbs}
emtrends(intmodel, ~RUBUS, "RUBUS", resp = "FORBCOVER", epred = TRUE)
```

Effects of rubus on grasses:

```{r bramble effect grasses}
emtrends(intmodel, ~RUBUS, "RUBUS", resp = "GRASSCOVER", epred = TRUE)
```

```{r}
(i6 <- plot(conditional_effects(intmodel, "RUBUS", resp = "FORBCOVER"),
           plot = FALSE)[[1]] +
    labs(x = "Bramble cover (%)", y = "Forb cover (%)"))
(i7 <- plot(conditional_effects(intmodel, "RUBUS", resp = "GRASSCOVER"),
           plot = FALSE)[[1]] +
    labs(x = "Bramble cover (%)", y = "Grass cover (%)"))
i1+i2+i4+i5+i3+i6+i7+plot_layout(nrow=2, guides = "collect") + plot_annotation(tag_levels = "a", tag_suffix = ")")
ggsave("Mult model deer and DB and rubus effects.png", path = "Outputs/",
       width = 15, height = 15, units = "cm", dpi = 300, scale= 1.5)
```



Effect of rubus on species richness:

```{r bramble effect sr number}
emtrends(intmodel, ~RUBUS, "RUBUS", resp = "SPECIESRICH", epred = TRUE)
```

Effect of forb on species richness:

```{r forb effect sr number}
emtrends(intmodel, ~FORB_COVER, "FORB_COVER", resp = "SPECIESRICH", epred = TRUE)
```

Effect of grasses on species richness:

```{r grass effect sr number}
emtrends(intmodel, ~GRASS_COVER, "GRASS_COVER", resp = "SPECIESRICH", epred = TRUE)
```

Effect of rubus on regen:

```{r bramble effect regen}
emtrends(intmodel, ~RUBUS, "RUBUS", resp = "REGEN", epred = TRUE)
```

Effect of forb on regen:

```{r forb effect regen}
emtrends(intmodel, ~FORB_COVER, "FORB_COVER", resp = "REGEN", epred = TRUE)
```

Effect of grasses on regen:

```{r grass effect regen}
emtrends(intmodel, ~GRASS_COVER, "GRASS_COVER", resp = "REGEN", epred = TRUE)
```

Plots of effects:

```{r bramble effect plot, fig.dim = c(7,5)}
cond <- data.frame(YR = "2022", ASHDIEBACK = "NoAsh", DEER = "Moderate")
p1 <- plot(conditional_effects(intmodel, "FORB_COVER", resp = "SPECIESRICH", 
                         conditions = cond), plot = FALSE)[[1]] + 
  labs(x = "Forb Cover (%)", y = "Ground Flora Richness")
p2 <- plot(conditional_effects(intmodel, "GRASS_COVER", resp = "SPECIESRICH", 
                         conditions = cond), plot = FALSE)[[1]] + 
  labs(x = "Grass Cover (%)", y = "Ground Flora Richness") 
p3 <- plot(conditional_effects(intmodel, "RUBUS", resp = "SPECIESRICH", 
                         conditions = cond), plot = FALSE)[[1]] +
  labs(x = "Bramble Cover (%)", y = "Ground Flora Richness")
p4 <- plot(conditional_effects(intmodel, "FORB_COVER", resp = "REGEN", 
                         conditions = cond), plot = FALSE,
           line_args = list(linetype = 2))[[1]] + 
  labs(x = "Forb Cover (%)", y = "Tree Regeneration") 
p5 <- plot(conditional_effects(intmodel, "GRASS_COVER", resp = "REGEN", 
                         conditions = cond), plot = FALSE)[[1]] + 
  labs(x = "Grass Cover (%)", y = "Tree Regeneration")
p6 <- plot(conditional_effects(intmodel, "RUBUS", resp = "REGEN", 
                         conditions = cond), plot = FALSE)[[1]] +
  labs(x = "Bramble Cover (%)", y = "Tree Regeneration")
p7 <- plot(conditional_effects(intmodel, "RUBUS", resp = "FORBCOVER", 
                         conditions = cond), plot = FALSE)[[1]] + 
  labs(x = "Bramble Cover (%)", y = "Forb Cover (%)")
p8 <- plot(conditional_effects(intmodel, "RUBUS", resp = "GRASSCOVER"),
           plot = FALSE)[[1]] + 
  labs(x = "Bramble Cover (%)", y = "Grass Cover (%)")
p1+p2+p3+p4+p5+p6+p7+p8 &
  theme(axis.title = element_text(size = 15))

p1 + p2 + p3 + p4 + p5 + p6 + plot_layout(axis_titles = "collect") &
  theme(axis.title = element_text(size = 15))
ggsave("Mult model sprich regen resp dash95ci.png", path = "Outputs/", 
       width = 15, height = 10, units = "cm", dpi = 600, scale = 1.5)
```


Climate effects:

```{r clim effect plot without data, fig.dim = c(8,8)}
p1 <- plot(conditional_effects(intmodel, "Winter_tasmin", resp = "FORBCOVER"),
           plot = FALSE, line_args = list(linetype = 2))[[1]] +
  labs(x = "Winter min temp (°C)", y = "Forb cover (%)") +
  scale_x_continuous(breaks = -3:2, labels = -2:3, limits = c(-3,2.5), expand = c(0,0)) 
p2 <- plot(conditional_effects(intmodel, "Winter_rainfall", resp = "FORBCOVER"),
           plot = FALSE, line_args = list(linetype = 2))[[1]] +
  labs(x = "Winter rainfall (mm)", y = "Forb cover (%)") +
  scale_x_continuous(breaks = seq(-2,6,2), labels = seq(100,900,200), limits = c(-2,7), expand = c(0,0))
p3 <- plot(conditional_effects(intmodel, "Winter_tasmin", resp = "RUBUS"),
           plot = FALSE)[[1]] +
  labs(x = "Winter min temp (°C)", y = "Bramble cover (%)") +
  scale_x_continuous(breaks = -3:2, labels = -2:3, limits = c(-3,2.5), expand = c(0,0)) 
p4 <- plot(conditional_effects(intmodel, "Winter_rainfall", resp = "RUBUS"),
           plot = FALSE)[[1]] +
  labs(x = "Winter rainfall (mm)", y = "Bramble cover (%)") +
  scale_x_continuous(breaks = seq(-2,6,2), labels = seq(100,900,200), limits = c(-2,7), expand = c(0,0)) 
p5 <- plot(conditional_effects(intmodel, "Winter_tasmin", resp = "SPECIESRICH"),
           plot = FALSE)[[1]] +
  labs(x = "Winter min temp (°C)", y = "Ground Flora richness") +
  scale_x_continuous(breaks = -3:2, labels = -2:3, limits = c(-3,2.5), expand = c(0,0)) 
p6 <- plot(conditional_effects(intmodel, "Winter_rainfall", resp = "SPECIESRICH"),
           plot = FALSE, line_args = list(linetype = 2))[[1]] +
  labs(x = "Winter rainfall (mm)", y = "Ground Flora richness") +
  scale_x_continuous(breaks = seq(-2,6,2), labels = seq(100,900,200), limits = c(-2,7), expand = c(0,0)) 
p7 <- plot(conditional_effects(intmodel, "Winter_tasmin", resp = "GRASSCOVER"),
           plot = FALSE)[[1]] +
  labs(x = "Winter min temp (°C)", y = "Grass cover (%)") +
  scale_x_continuous(breaks = -3:2, labels = -2:3, limits = c(-3,2.5), expand = c(0,0)) 
p8 <- plot(conditional_effects(intmodel, "Winter_rainfall", resp = "GRASSCOVER"),
           plot = FALSE, line_args = list(linetype = 2))[[1]] +
  labs(x = "Winter rainfall (mm)", y = "Grass cover (%)") +
  scale_x_continuous(breaks = seq(-2,6,2), labels = seq(100,900,200), limits = c(-2,7), expand = c(0,0)) 
p9 <- plot(conditional_effects(intmodel, "Winter_tasmin", resp = "REGEN"),
           plot = FALSE, line_args = list(linetype = 2))[[1]] +
  labs(x = "Winter min temp (°C)", y = "Tree regeneration") +
  scale_x_continuous(breaks = -3:2, labels = -2:3, limits = c(-3,2.5), expand = c(0,0)) 
p10 <- plot(conditional_effects(intmodel, "Winter_rainfall", resp = "REGEN"),
            plot = FALSE)[[1]] +
  labs(x = "Winter rainfall (mm)", y = "Tree regeneration") +
  scale_x_continuous(breaks = seq(-2,6,2), labels = seq(100,900,200), limits = c(-2,7), expand = c(0,0)) 


p3 + p4 + p1 + p2 + p7 + p8 + p5 + p6 + p9 + p10 +plot_layout(nrow = 5, axis_titles = "collect")
```

```{r}
ydays <- (c(121,152,182,213,244)-200)/30
yday_labs <- paste0("01-",month.abb[5:9])
y1 <- plot(conditional_effects(intmodel, "YDAY", resp = "RUBUS"),
            plot = FALSE)[[1]] +
  labs(x = "Day of year", y = "Bramble cover (%)") +
  scale_x_continuous(breaks = ydays, labels = yday_labs)
y2 <- plot(conditional_effects(intmodel, "YDAY", resp = "FORBCOVER"),
            plot = FALSE)[[1]] +
  labs(x = "Day of year", y = "Forb cover (%)") +
  scale_x_continuous(breaks = ydays, labels = yday_labs)
y3 <- plot(conditional_effects(intmodel, "YDAY", resp = "GRASSCOVER"),
            plot = FALSE)[[1]] +
  labs(x = "Day of year", y = "Grass cover (%)") +
  scale_x_continuous(breaks = ydays, labels = yday_labs)
y4 <- plot(conditional_effects(intmodel, "YDAY", resp = "SPECIESRICH"),
            plot = FALSE)[[1]] +
  labs(x = "Day of year", y = "Ground Flora richness") +
  scale_x_continuous(breaks = ydays, labels = yday_labs)
y5 <- plot(conditional_effects(intmodel, "YDAY", resp = "REGEN"),
            plot = FALSE, line_args = list(linetype = 2))[[1]] +
  labs(x = "Day of year", y = "Tree regeneration") +
  scale_x_continuous(breaks = ydays, labels = yday_labs)

p3 + p4 + y1 + p1 + p2 + y2 + p7 + p8 + y3 + p5 + p6 + y4 + p9 + p10 + y5 +
  plot_layout(nrow = 5, axis_titles = "collect")
ggsave("Mult model clim yday effects dash95ci.png", path = "Outputs/",
       width = 15, height = 15, units = "cm", scale = 1.5)
```


Residual correlation:

```{r}
total_res <- resid(intmodel)

comb_res <- data.frame(SPECIESRICH = total_res[,1,"SPECIESRICH"],
                       REGEN = total_res[,1,"REGEN"],
                       FORBCOVER = total_res[,1,"FORBCOVER"],
                       GRASSCOVER = total_res[,1,"GRASSCOVER"],
                       RUBUS = total_res[,1,"RUBUS"])
```

```{r}
psych::pairs.panels(comb_res, global = FALSE, method = "spearman")
```

```{r}
ggplot(comb_res, aes(x = REGEN, y = SPECIESRICH)) + 
  geom_point() +
  geom_smooth()
```

